# Reproducible Research: Peer Assessment 1
```{r, results='hide'}
#load packages
library("ggplot2")

#get english localization settings
Sys.setlocale("LC_TIME", "en_GB")
```
Make sure the `activity.csv` file is in your working directory.


<br><br>
## Loading and preprocessing the data
```{r}
#load data
activity <- read.csv("activity.csv")

#data column from factor to date class
activity$date <- as.Date(activity$date)

```

<br><br>
## What is mean total number of steps taken per day?
```{r}
#sum the number of steps per day
total_steps <- tapply(activity$steps, activity$date, sum)

hist(total_steps, col="blue", main = "Total number of steps taken each day", xlab = " Total number of steps")
```

_Figure 1: Histogram of the total number of steps taken each day. On the x-axis you see the number of steps, and on the y-axis the number of days_


```{r}
mean(total_steps, na.rm=TRUE)
median(total_steps, na.rm=TRUE)

```

The mean total number of steps taken per day is `10766` and the median total number of steps taken per day is `10765`.

<br><br>
## What is the average daily activity pattern?
```{r}
average_steps <- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)

plot(average_steps, type="l", col="blue", main = "Average daily activity pattern", 
     xlab= "Time", ylab="Average number of steps" )
```

_Figure 2: Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)._


```{r}
names(average_steps[which.max(average_steps)])
```

The 5-minute interval that contains the maximum number of steps on average across all the days in the dataset is interval `835`.

<br><br>
## Imputing missing values

```{r}
sum(is.na(activity))
```

The total number of missing values in the dataset is `2304`.

**Strategy for filling in all of the missing values in the dataset:**
* Create an extra column that contains the averge steps per 5 min interval
* If the number of steps is `NA` replace the value with the value in the averege_steps columns

```{r}
#Create a new dataset that is equal to the original dataset but with the missing data filled in.
d<-activity
d$average_steps <- average_steps
d[is.na(d),]$steps <- d[is.na(d),]$average_steps
```

```{r}
total_steps2 <- tapply(d$steps, d$date, sum)

hist(total_steps2, col="blue", main = "Total number of steps taken each day", xlab = " Total number of steps")
```

_Figure 3: Histogram of the total number of steps taken each day after filling in the missing values. On the x-axis you see the number of steps, and on the y-axis the number of days._

```{r}
mean(total_steps2)
median(total_steps2)
```
The mean total number of steps taken per day is `10766` and the median total number of steps taken per day is `10766`.

After filling in the missing values the mean remains the same, however the median is now identical to the mean. 

<br><br>
## Are there differences in activity patterns between weekdays and weekends?

```{r}
#Create a new factor variable in the dataset with two levels  "weekday" and "weekend" 
d$day <- as.factor(weekdays(d$date))

levels(d$day)<- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),
                     weekend = c("Sunday", "Saturday"))

```

```{r}
ggplot(data=d, aes(interval,steps)) + facet_grid(day~.) + 
        stat_summary(fun.y=mean, col="blue", geom="line") +
        xlab("Interval") + ylab("Average number of steps") +
        ggtitle("Averaged steps across all weekday days or weekend days")

```

_Figure 4: Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)._


<br><br>
## References
* H. Wickham. ggplot2: elegant graphics for data analysis. Springer New York, 2009.
* Yihui Xie (2014). knitr: A general-purpose package for dynamic report generation in R. R package version 1.6.
* Yihui Xie (2013) Dynamic Documents with R and knitr. Chapman and Hall/CRC. ISBN 978-1482203530


<br><br>
## Session Info
```{r}
sessionInfo()

```
